Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-03-02 resex.tex
Another intermediate step in the revision of the results section: discussion and plot of calculated beam re-distribution.

\begin{figure*}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$10^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		random H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_3_5um_hist_y.png}
	\end{minipage}	
	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn pulsing (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from distribution tracking simulations based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ and standard errors. The excitation is applied in both planes with an amplitude of 96~nrad. The residual is defined as the final distribution (here after $10^4$~turns) minus the initial distribution (here after $10^2$~turns) as \%, and the ratio as the final distribution divided by the initial distribution. The red line in the plot of the residual is the difference between the Gaussian fit of the distribution and the distribution itself.}
\end{figure*}

@@

One may have noticed that the emittance in case of the
$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an
increased initial value and then stays almost constant during the
entire simulation time, which usually points to an error in the chosen
simulation setup. This behavior is however typical for any resonant
excitation and can be ascribed to the adjustment of the beam
distribution during the first $10^4$ turns to a new equilibrium. As
the beam distribution is only dumped every $10^4$ turns, this initial
fast adjustment manifests itself in an increased initial emittance
value in Fig.~\ref{fig:patternsim}, and also any other simulations
over $10^6$ turns presented in this paper. Besides an increased
emittance, the new distribution also becomes non-Gaussian. This is
illustrated by means of the residual in respect to the Gaussian
distribution for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing
in Fig.~\ref{fig:patternhist}. This initial fast change of the beam
distribution is not an artifact of the simulation, but has been, to
the knowledge of the authors, also for the first time experimentally
observed during the 2016 and 2017 LHC experiments presented in this
paper. The timescale of this fast adjustment is in experiments however
much slower than predicted in simulations --- on the timescale of
minutes compared to seconds.

>>>>

\begin{figure*}
  \begin{tabular}{cc}
    no excitation & random, H+V \\
    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_3_5um_hist_y.png} &
    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
    $7^{\mathrm{th}}$, H+V & $10^{\mathrm{th}}$, H+V \\
    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
                  &
    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
  \end{tabular}
  \caption{Calculated beam distributions as a function of vertical
    position from distribution-tracking simulations based on the 2016
    injection optics with $(Q_x, Q_y) = (64.28, 59.31)$ and standard
    lattice errors: no excitation (top left), random excitation (top
    right), $7^{\mathrm{th}}$~turn pulsing (bottom left), and
    $10^{\mathrm{th}}$~turn pulsing (bottom right). The excitations
    are applied in both planes with an amplitude of 96~nrad. For each
    of the 4~cases, 3~plots are shown. The top plot shows the
    normalized transverse distributions: `initial' (after
    $10^2$~turns, in gray), `final' (after $10^4$~turns, in black),
    and their Gaussian fits (light and dark red, respectively). The
    middle plots show the relative residuals (i.e., differences, in
    percent of the peak value) between final and initial distributions
    (in black) and between each distribution and its Gaussian fit (in
    light and dark red). The ratios between final and initial
    distributions are drawn in black in the bottom plots. The
    residuals emphasize changes near the core of the distributions,
    whereas ratios (when statistically significant), reveal variations
    in the tails.}
  \label{fig:patternhist}
\end{figure*}

@@

In Fig.~\ref{fig:patternsim}, one can see that, in the cases of
$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing, the emittances
start from an increased initial value and then stay almost constant
for the duration of the simulation. This behavior is typical of the
resonant excitation and it is not an artifact of the simulation. It
was verified that it is due to the adjustment of the input beam
distribution to a new equilibrium during the first $10^4$~turns. In
all long-term simulations ($10^6$~turns, corresponding to 89~s)
presented in this paper, the beam distribution is saved every
$10^4$~turns. Therefore, this initial fast adjustment manifests itself
in an increased initial emittance value.

Besides having a larger emittance, the new distribution also differs
from a Gaussian distribution.  This is illustrated in
Fig.~\ref{fig:patternhist}, which shows, for different excitation
patterns, two snapshots of the normalized transverse beam
distributions, the `initial' one at $10^2$~turns and the `final' one
at $10^4$~turns, and how they differ from each other and from a
Gaussian distribution. For instance, for $7^{\mathrm{th}}$ turn
pulsing (bottom left plot in Fig.~\ref{fig:patternhist}), simulations
indicate a clear shift in beam population from the core ($< 1\sigma$)
towards the regions at $\pm 1.5\sigma$ and towards the tails at
$\pm 4\sigma$.

Although on a different time scale (minutes instead of seconds), a
similar re-adjustment of the beam distribution in response to an
external excitation was directly observed experimentally for the first
time, to our knowledge, during the 2016 and 2017 LHC experiments
presented in this paper, as discussed below.
